---
title: "R Notebook"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

JSM is Largest statistical event. This report is for Data Expo 2018.
There are 3 data files, locations.csv, forecast.dat, and histWeather.csv.
1- The locations.csv file is a comma separated value file that contains information on the cities for which the forecasts was made
2- The forecast.dat file is a white space separated file with about 3 years worth of forecasts.  This file does not have a header row.  The first column is the city number corresponding to the row in the locations.csv file, so 1 means Eastport, Maine and 113 means Honolulu, Hawaii.

3- The histWeather.csv file is a comma separated file with the historic measures of weather from the airports.

 {r} [] (http://community.amstat.org/stat-computing/data-expo/data-expo-2018)
 



```{r}
#Imporing Forecast datafile
library(readr)
data1 <- read_csv("C:/Users/chink/Google Drive/Computational Statistics/5th Quater/Special Topics/JSM/forecast.dat", col_names = FALSE)
```

```{r}
#Converting datafiles into columns 
data1$a <- sapply(strsplit(as.character(data1$X1), " "), "[", 1)
data1$b <- sapply(strsplit(as.character(data1$X1), " "), "[", 2)
data1$c<- sapply(strsplit(as.character(data1$X1), " "), "[", 3)
data1$d<- sapply(strsplit(as.character(data1$X1), " "), "[", 4)
data1$e<- sapply(strsplit(as.character(data1$X1), " "), "[", 5)
data1$c=as.numeric(data1$c)

#Structure of dataframe
str(data1)

```

```{r}
#Reading Histweather data
histWeather <- read_csv("C:/Users/chink/Google Drive/Computational Statistics/5th Quater/Special Topics/JSM/histWeather.csv")
histWeather$CityCode=as.factor(histWeather$CityCode)
histWeather$PrecipitationIn=as.integer(histWeather$PrecipitationIn)
histWeather$AirPtCd=as.factor(histWeather$AirPtCd)

#Structure of histweather data
str(histWeather)
```
```{r}
#Summary of variables
summary(histWeather)
```
Max_TemperatureF has max value 211119.00.Which seems an outlier.
```{r}
#Finding the index of max variable
which(histWeather$Max_TemperatureF==211119.00)
```

```{r}
#Making month,year and day 
library(lubridate)
class(histWeather$Date)
histWeather$dat=mdy(histWeather$Date)
class(histWeather$dat)
histWeather$month=month(histWeather$dat)
histWeather$day=day(histWeather$dat)
histWeather$year=year(histWeather$dat)
```

```{r}
#Ploting Min_Temp
library(ggplot2)
ggplot(aes(x=Date,y=Min_TemperatureF),data=histWeather)+geom_line(col="green", size=1)
```

```{r}
#ggplot of Min_TeamperaturF and month
ggplot(aes(x=month,y=Min_TemperatureF),data=histWeather)+geom_point(col="green", size=1)+ylim(c(-150,150))+xlim(c(1,12))
```



```{r}
#Summary of data1
summary(data1)
str(data1)
data1$b=as.factor(data1$b)
data1$d=as.factor(data1$d)
str(data1)
```


```{r}
#Tidy data
#library(tidyr)
data1=data1[c(2,3,5,4)]
data1=as.data.frame(data1)
summary(data1$c)
data1[is.na(data1)]=0
dim(data1)
summary(data1$c)
library(dplyr)
library(tidyr)

#Convertings row to column
data2=data1 %>% group_by(d,b,a) %>% summarise_at(.vars = names(.)[4],.funs = c(mean="mean"))
data3=data2 %>% spread(d,mean)



write.table(data3, "C:/Users/chink/Google Drive/Computational Statistics/5th Quater/Special Topics/JSM/data3.txt", sep="\t")

```
```{r}
#Renaming columns 
names(data3)[names(data3)=='a']='CityCode'
names(data3)[names(data3)=='b']='Date'
names(data3)[names(data3)=='MinTemp']='P_MinTemp'
names(data3)[names(data3)=='MaxTemp']='P_MaxTemp'
names(data3)[names(data3)=='ProbPrecip']='P_ProbPrecip'

```


```{r}
#Changing formate of histweather data date col
histWeather$Date=as.Date(histWeather$Date,format = "%m/%d/%Y")

#Creating final data with merging of hisweather and forcasted data
library(plyr)
final=merge(histWeather,data3,by=c("CityCode","Date"))
```


```{r}
#Creating text file
library(readxl)
Prediction_work <- read_excel("C:/Users/chink/Google Drive/Computational Statistics/5th Quater/Special Topics/JSM/Prediction_work.xlsx")
```

```{r}
#Dufference of Max temp actual and predicted value
dif_Max_Temp=final$P_MaxTemp-final$Max_TemperatureF
dif_Max_Temp=as.data.frame(dif_Max_Temp)
summary(dif_Max_Temp)
```
```{r}
#Predicted vrs Actaul 
library(ggplot2)
ggplot(final,aes(x=P_MaxTemp,y=Max_TemperatureF,col="red"))+geom_point()+xlim(-25,125)+ylim(0,100)+xlab("Predicted Max Temp")+ylab("Actual Max Temp")

```
```{r}
#ploting diff and time
final$dif11=final$Max_TemperatureF-final$P_MaxTemp
ggplot(final,aes(x=Date,y=dif11,col="red"))+geom_line()+ylim(0,75)+xlab("Date")+ylab("Difference between predicted and Actual")
```

```{r}
#ploting diff and time
final$dif21=final$Min_TemperatureF-final$P_MinTemp
ggplot(final,aes(x=Date,y=dif21,col="red"))+geom_line()+ylim(0,50)+xlab("Date")+ylab("Difference between predicted and Actual")
```

```{r}
#ploting diff and time
final$dif31=as.numeric(final$PrecipitationIn)-final$P_ProbPrecip
ggplot(final,aes(x=Date,y=dif31,col="red"))+geom_line()+ylim(0,10)+xlab("Date")+ylab("Difference between predicted and Actual")
```

```{r}
#Absolute Error between predicted and actual values of Max Temp
final$dif12=abs(final$Max_TemperatureF-final$P_MaxTemp)
final$dif22=abs(final$Min_TemperatureF-final$P_MinTemp)
final$dif32=abs(as.numeric(final$PrecipitationIn)-final$P_ProbPrecip)


#Square Error between predicted and actual values of Max Temp
final$dif13=(final$Max_TemperatureF-final$P_MaxTemp)^2
final$dif23=(final$Min_TemperatureF-final$P_MinTemp)^2
final$dif33=(as.numeric(final$PrecipitationIn)-final$P_ProbPrecip)^2

#Summary Statistics
summary(final)

```

```{r}
#MAE of temp
MAE_Max_Temp=5.10
MAE_Min_Temp=3.831
MAE_prec=13.253

#MSE of temp
MSE_Max_Temp=3.589e+05 
MSE_Min_Temp=27.314 
MSE_prec=456.14 
```
MAE and MSE is higher for Max Temp becauses Max Temp has outliers.

```{r}
write.table(final, "C:/Users/chink/Google Drive/Computational Statistics/5th Quater/Special Topics/JSM/final.txt", sep="\t")
```

Lets try to find the distribution of errors 
```{r}
#"Histogram of Error of MaxTemp in F"
library(ggplot2)
ggplot(final,aes(dif11))+geom_histogram(col="red",fill='green')+xlim(c(-15,15))+xlab("Error of Max temp")
```

Plot is looking symmetric and bell shaped. normal distribution is justified distribution of error in Max Temp in F.

```{r}
#"Histogram of Error of MinTemp in F"
ggplot(final,aes(dif21))+geom_histogram(col="red",fill='green')+xlim(c(-25,25))+xlab("Error of Min temp")
```
Plot is looking symmetric and bell shaped. normal distribution is justified distribution of error in Min Temp in F.

```{r}
#"Histogram of Error of MinTemp in F"
ggplot(final,aes(dif31))+geom_histogram(col="red",fill='green')+xlim(c(-100,5))+xlab("Error of Min temp")
```
```{r}
#Normal qq plot 
qqnorm(final$dif31,col="green")
```
```{r}
#Normalizing the data
 normalize <- function(x) 
  { 
     return ((x - min(x)) / (max(x) - min(x)))
  }
#Removing NAs from data
diff <- final$dif31[!is.na(final$dif31)]

#Normalizing precip
x=normalize(diff)

summary(x)
```

```{r}
#Density plot of precip
plot(density(x),col="red")
```



![Distributions](https://magesblog.com/post/2011-12-01-fitting-distributions-with-r/)

I think Minimum extrem is justified distribution for Error in precip.



