---
title: "Titanic"
author: "Chinki"
date: "June 6, 2017"
output: word_document
---
Titanic Machine learning from Disaster
Data Collection: source-https://www.kaggle.com/c/titanic/data Data has two group training and test. Data description Variable Definition Key survival Survival 0 = No, 1 = Yes pclass Ticket class 1 = 1st, 2 = 2nd, 3 = 3rd sex Sex Age Age in years
sibsp # of siblings / spouses aboard the Titanic
parch # of parents / children aboard the Titanic
ticket Ticket number
fare Passenger fare
cabin Cabin number
embarked Port of Embarkation C = Cherbourg, Q = Queenstown, S = Southampton

I have two set of data. 
```{r}
#Reading the data
training=read.csv("train.csv",na.strings = T)
test=read.csv("test.csv",na.strings = T)
```
```{r}
#Table of pclass 
table(training$Sex,training$Survived,training$Pclass)
```
```{r}
#Table of pclass and sex
table(training$Pclass,training$Sex)
```
```{r}
#Boxplot of age
barplot(prop.table(table(training$Sex)),col = "blue")
```
```{r}
barplot(prop.table(table(training$Survived)))
```

```{r}
str(training)
```
```{r}
str(test)
```
```{r}
#Creating levels 
training$Survived=factor(training$Survived,levels = c("0","1"),labels = c("survived","notsurvived"))
```
```{r}
#checking for missing values 
sapply(training,function(x) sum(is.na(x)))
```
```{r}
#creating dummy Variables
contrasts(training$Embarked)
```
```{r}
#creating dummy Variables
contrasts(test$Embarked)
```

```{r}
#Creating dummy variables 
contrasts(training$Sex)
```
```{r}
#Creating dummy variables 
contrasts(test$Sex)
```

```{r}
#In tarining data set 177 missing values
training$Age[is.na(training$Age)]=mean(training$Age,na.rm = T)
test$Age[is.na(test$Age)]=mean(test$Age,na.rm = T)
```
```{r}
train=training[c(2,3,5,6,7,8,10,12)]
```
```{r}
#Logistic Regression model
result=glm(Survived~.,family = binomial(link = "logit"),data = train,control = list(maxit = 1500))
summary(result)
```
```{r}
anova(result,test="Chisq")
```
Pclass , sex , age and sibsp is the significant variable.
```{r}
#Test data selection
t=test[c(2,4,5,6,7,9,11)]
```
```{r}
#Fitted value
fitted_result=predict(result,t,type="response")
```
```{r}
#Checking for survived or not survived 
fitted=ifelse(fitted_result>0.5,1,0)
output=data.frame(test$PassengerId,fitted)
```
```{r}
library(xlsx)
write.xlsx(output,file="output.xlsx")

```





